---
title: "R Notebook"
output: html_notebook
---

# Install Packages 

```{r}
#install.packages("EpiModel")
library(EpiModel)
```

# Basic SI Model Tutorial

## Model Building

```{r}
param <- param.dcm(inf.prob = 0.2, act.rate= 0.25) # parameters
init <- init.dcm(s.num = 500, i.num = 1) # collects initial conditions
control <- control.dcm(type="SI", nsteps = 500) # collects structural model controls 

model <-dcm(param,init,control) # run the model
model
```
two compartments and one flow are available. Compartments are discrete disease states and flows are the transitions between states. The i.num compartment is the size of the infected population at each of the time steps. si.flow is the number of people movng from S to I at each time step. i.num is disease prevalence and si.flow is disease incidence. 

## Plot

```{r}
plot(model)
```

## Summary

```{r}
summary(model, at=150)
```

Disease incidence is 4.31 new infections

# SIR Model with Demography Tutorial

## Model

```{r}
param.sir <- param.dcm(inf.prob = 0.2, act.rate=1, rec.rate=1/20, a.rate = 1/95,ds.rate=1/100,di.rate=1/80,dr.rate = 1/100)

init.sir <- init.dcm(s.num = 1000, i.num = 1, r.num = 0)
control.sir <- control.dcm(type="SIR", nsteps = 500, dt=0.5)

model.sir <- dcm(param.sir,init.sir,control.sir)
model.sir
```

## Plot

```{r}
par(mar = c(3.2,3,2,1), mgp=c(2,1,0), mfrow=c(1,2))
plot(model.sir, popfrac = FALSE, alpha = 0.5, lwd = 4, main= "Compartment Sizes")
plot(model.sir, y= "si.flow", lwd = 4, col = "firebrick", main = "Disease Incidence", legend="n")

```

## Summaries

```{r}
par(mfrow = c(1,1))
comp_plot(model.sir, at= 50, digits=1)
```

# Basic ICMs Tutorial

## SI Model Stochasticity

```{r}
param.si.icm <- param.icm(inf.prob = 0.2, act.rate = 1, rec.rate = 1/20)
init.si.icm <- init.icm(s.num = 500, i.num = 1, r.num = 0)
control.si.icm <- control.icm(type="SIR", nsims = 10 ,nsteps = 300)
model.si.icm <- icm(param.si.icm,init.si.icm, control.si.icm)
model.si.icm
```

## Summaries

```{r}
summary(model.si.icm, at= 125)
```

```{r}
head(as.data.frame(model.si.icm, out="mean"))
```

```{r}
tail(as.data.frame(model.si.icm, out="vals", sim=1))
```

## Plotting

```{r}
plot(model.si.icm)
```

# Problem 4: Investigsting Simulated Outbreaks 

## DCM

### Model

```{r}

param.ny <- param.dcm(inf.prob = 0.7, act.rate=2.2, rec.rate=1/14, a.rate = 0.014,ds.rate=0, di.rate=0.2, dr.rate = 0)

param.ny <- param.dcm(inf.prob = 0.7, act.rate=2.2, rec.rate=1/14, a.rate = 0.014,ds.rate=0,di.rate=0.2,dr.rate = 0)


init.ny <- init.dcm(s.num = 8399000, i.num = 1, r.num = 0)
control.ny <- control.dcm(type="SIR", nsteps = 500, dt=0.5)

model.ny <- dcm(param.ny,init.ny,control.ny)
model.ny
```

### Plot

```{r}
par(mar = c(3.2,3,2,1), mgp=c(3,1,0), mfrow=c(1,2))
plot(model.ny, popfrac = FALSE, alpha = 0.5, lwd = 4, main= "Compartment Sizes")

plot(model.ny, y = "si.flow", lwd = 4, col = "firebrick", main = "Disease Incidence", legend="n")
```

## Summaries

```{r}
par(mfrow = c(1,1))
comp_plot(model.ny, at= 50, digits=1)

plot(model.ny, y= "si.flow", lwd = 4, col = "firebrick", main = "Disease Incidence", legend="n", ylab="SI Flow")
```

### Summary 

```{r}
summary(model.ny, at= 150)
```

# ICM

## Model

```{r}
param.icm.ny <- param.icm(inf.prob = 0.7,act.rate=2.2, rec.rate=1/14, a.rate = 0.014,ds.rate=0,di.rate=0.2,dr.rate = 0)
init.icm.ny <- init.icm(s.num = 124000, i.num = 1, r.num = 0)
control.icm.ny <- control.icm(type="SIR", nsims = 2,nsteps = 300)

model.icm.ny <- icm(param.icm.ny,init.icm.ny, control.icm.ny)
model.icm.ny
```

## Plot

```{r}
plot(model.icm.ny)
```

## Summary

```{r}
summary(model.icm.ny, at=125)
```


# Problem 5

```{r}
param.icm.ny.2 <- param.icm(inf.prob = 0.7,act.rate=2.2, rec.rate=1/14, a.rate = 0.014,ds.rate=0,di.rate=0.2,dr.rate = 0)
init.icm.ny.2 <- init.icm(s.num = 124000, i.num = 1, r.num = 0)
control.icm.ny.2 <- control.icm(type="SIR", nsims = 2,nsteps = 300) 

model.icm.ny.2 <- icm(param.icm.ny.2,init.icm.ny.2, control.icm.ny.2)
model.icm.ny.2
```


# Problem 6

Change the act rate to be low, since quarantine prevents interaction. Additionally, decrease the probability of infection since quarantine reduces the risk. 

```{r}
param.ny.intervention <- param.dcm(inf.prob = 0.5, act.rate=1, rec.rate=1/14, a.rate = 0.014,ds.rate=0,di.rate=0.2,dr.rate = 0)

init.ny.intervention <- init.dcm(s.num = 8399000, i.num = 1, r.num = 0)
control.ny.intervention <- control.dcm(type="SIR", nsteps = 500, dt=0.5)

model.ny.intervention <- dcm(param.ny.intervention,init.ny.intervention,control.ny.intervention)
model.ny.intervention
```

```{r}
plot(model.ny.intervention)
```

```{r}
summary(model.ny.intervention, at=150)
```

These samples are much smaller than the first DCM model because quarantine decreases the probability of infection and the act rate. The number of susceptibles is higher and infected is lower. SI is also smaller, because incident rate is lower. 
